Valuing Real Estate in Seattle Using Property Characteristics: A Machine Learning Approach

### 2024-11-28 paper/paper.qmd
Adding charts

library(tidyverse)
library(palmerpenguins)
library(kableExtra)
library(arrow)
library(stargazer)
library(ggpubr)
library(modelsummary)
library(dagitty)
library(ggdag)
```

@@

```{r fig-out, fig.width=7, fig.height=2.5, fig.cap='Heavily Skewed Property Prices and Skewed Still for Logarithmic Prices'}
p1 <- data |>
  select(price) |>
  ggplot(aes(x = price)) +
  geom_histogram(binwidth = 1000000, fill = "beige", color = "black") +
  labs(x = "Price ($)", y = "Count")

p2 <- data |>
  select(price) |>
  ggplot(aes(x = log(price))) +
  geom_histogram(binwidth = 0.25, fill = "beige", color = "black") +
  labs(x = " ln(Price)", y = "Count")

@@

```{r fig-pred1, fig.width=7, fig.height=2.5, fig.cap='Single Family Homes and 2 or 3 Bedrooms Properties Most Listing on the Market'}
p3 <- data |> 
  ggplot(aes(x = property_type)) +
  geom_bar(fill = "beige", color = "black") +
  labs(x = "Property Type", y = "Count")

p4 <- data |> 
  mutate(beds = as.factor(beds)) |>
  ggplot(aes(x = beds)) +
  geom_bar(fill = "beige", color = "black") +
  labs(x = "Beds", y = "Count")

@@

```{r fig-pred2, fig.width=7, fig.height=2.5, fig.cap='Skewed Distribution of Sizes and Most Properties Built Over the Last Decade'}
p5 <- data |>
  ggplot(aes(x = sqft)) + 
  geom_histogram(binwidth = 500, fill = "beige", color = "black") +
  labs(x = "Size (sqft)", y = NULL)

p6 <- data |> 
  ggplot(aes(x = property_age)) + 
  geom_histogram(binwidth = 10, fill = "beige", color = "black") +
  labs(x = "Property Age (years)", y = NULL)

@@

```{r fig-dag, fig.width=4, fig.height=3, fig.cap='Multiple Causal Relationships Among Model Variables'}
dagify(
  price ~ beds,
  price ~ sqft,
  price ~ baths,
  price ~ hoa,
  sqft ~ beds,
  sqft ~ baths,
  baths ~ beds,
  hoa ~ sqft,
  coords = list(
    x = c(beds = 0.5, sqft = 1, baths = 0, price = 2, hoa = 1.5),
    y = c(beds = 0, sqft = 1, baths = 1, price = 0, hoa = 1.5)
  )
) |>
  ggdag() +
  theme_void()
```

@@

@@

```{r, fig.align='center'}
palmerpenguins::penguins |>
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm, color = species)) +
  geom_point()
```

@@

```{r, fig.align='center'}
palmerpenguins::penguins |>
  drop_na() |>
  ggplot(aes(x = sex, y = body_mass_g, fill = sex)) +
  geom_boxplot(width = 0.5) +
  coord_flip()
```

@@

```{r, fig.align='center'}
palmerpenguins::penguins |>
  drop_na() |>
  ggplot(aes(x = species, y = flipper_length_mm, fill = species)) +
  geom_violin(alpha = 0.5) +
  geom_boxplot(width = 0.25) +
  coord_flip()
```

>>>>

library(tidyverse)
library(rsample)
library(kableExtra)
library(arrow)
library(stargazer)
library(ggpubr)
library(modelsummary)
library(dagitty)
library(ggdag)
```

@@

```{r fig-out, fig.width=7, fig.height=2.5, fig.cap='Heavily Skewed Property Prices and Skewed Still for Logarithmic Prices'}
p1 <- data |>
  select(price) |>
  ggplot(aes(x = price)) +
  geom_histogram(binwidth = 1000000, fill = "#519C7C", color = "black") +
  labs(x = "Price ($)", y = "Count")

p2 <- data |>
  select(price) |>
  ggplot(aes(x = log(price))) +
  geom_histogram(binwidth = 0.25, fill = "#519C7C", color = "black") +
  labs(x = " ln(Price)", y = "Count")

@@

```{r fig-pred1, fig.width=7, fig.height=2.5, fig.cap='Single Family Homes and 2 or 3 Bedrooms Properties Most Listing on the Market'}
p3 <- data |> 
  ggplot(aes(x = property_type)) +
  geom_bar(fill = "#519C7C", color = "black") +
  labs(x = "Property Type", y = "Count")

p4 <- data |> 
  mutate(beds = as.factor(beds)) |>
  ggplot(aes(x = beds)) +
  geom_bar(fill = "#519C7C", color = "black") +
  labs(x = "Beds", y = "Count")

@@

```{r fig-pred2, fig.width=7, fig.height=2.5, fig.cap='Skewed Distribution of Sizes and Most Properties Built Over the Last Decade'}
p5 <- data |>
  ggplot(aes(x = sqft)) + 
  geom_histogram(binwidth = 500, fill = "#519C7C", color = "black") +
  labs(x = "Size (sqft)", y = NULL)

p6 <- data |> 
  ggplot(aes(x = property_age)) + 
  geom_histogram(binwidth = 10, fill = "#519C7C", color = "black") +
  labs(x = "Property Age (years)", y = NULL)

@@

```{r fig-dag, fig.align='center', fig.width=4, fig.height=3, fig.cap='Multiple Causal Relationships Among Model Variables'}
dagify(
  price ~ beds,
  price ~ sqft,
  price ~ baths,
  price ~ hoa,
  sqft ~ beds,
  sqft ~ baths,
  baths ~ beds,
  hoa ~ sqft,
  coords = list(
    x = c(beds = 0.5, sqft = 1, baths = 0, price = 2, hoa = 1.5),
    y = c(beds = 0, sqft = 1, baths = 1, price = 0, hoa = 1.5)
  )
) |>
  ggdag() +
  theme_void()
```

@@

```{r}
set.seed(12)

split <- initial_split(data, prop = 0.7)

testing_data <- testing(split)
```

@@

```{r fig-scatter, fig.cap='Large Uncertainty in the Values of Luxury Homes and Mansions'}
testing_data |>
  mutate(predicted_price = predict(final_model, newdata = testing_data)) |>
  drop_na() |>
  ggplot(aes(x = predicted_price / 10^6, y = price / 10^6)) +
  geom_point(color = "#519C7C", size = 1) +
  geom_function(fun = function(x){x}, linetype = "dashed") +
  scale_x_continuous(breaks = seq(0, 20, 2)) +
  scale_y_continuous(breaks = seq(0, 20, 2)) +
  labs(x = "Predicted Price ($ millions)", y = "Actual Price ($ millions)")
```

@@

```{r fig-col, fig.cap='Highest Average Predicted Listing Price in 85 Acre Gated Community'}
testing_data |>
  mutate(predicted_price = predict(final_model, newdata = testing_data)) |>
  drop_na() |>
  select(neighbourhood, predicted_price) |>
  group_by(neighbourhood) |>
  summarise(mean = mean(predicted_price / 10^6)) |>
  arrange(desc(mean)) |>
  head(n = 5) |>
  ggplot(aes(x = reorder(neighbourhood, mean), y = mean)) +
  geom_col(position = "dodge", fill = "#519C7C", color = "black") +
  scale_y_continuous(breaks = seq(0, 5, 0.5)) +
  coord_flip() +
  labs(x = "Neighbourhood", y = "Mean Predicted Price ($ millions)")
```

@@

```{r fig-box, fig.cap='Single Family Homes Valued Significantly Higher Than Other Property Types'}
testing_data |>
  mutate(predicted_price = predict(final_model, newdata = testing_data)) |>
  drop_na() |>
  ggplot(aes(x = as.character(property_type), y = predicted_price / 10^6)) +
  geom_boxplot(fill = "#519C7C", width = 0.25) +
  coord_flip() +
  scale_y_continuous(breaks = seq(0, 8, 1)) +
  labs(x = "Property Type", y = "Price ($ millions)")
```
